Prey movement shapes the development of predator expertise in a virtual bi-trophic system

### 2024-08-16 manuscript/manuscript.Rmd
updated abstract

The acquisition of expertise is crucial for predators to be successful hunters. To do so, predators need to hone their skills and acquire knowledge through repeated and extensive practice. Recent studies suggest that such acquisition of expertise may not follow expected patterns when prey use antipredator tactics to evade detection and pursuit. Such tactics may even impair the acquisition of predator expertise. Yet, there are limited empirical assessments showing how predators acquire expertise through repeated encounters with their prey due to the challenges of monitoring direct interactions in the wild. Here, we use a virtual predator-prey system (the game *Dead by Daylight*) to investigate how experience shapes individual and population hunting success in human predators across repeated interactions with their prey. We show that predators optimized prey consumption as they gained experience, indicating that they acquired expertise through extensive practice. Prey speed was an important mediator of this relationship, driving differences among predators in the development of expertise. At the population-level, we found that faster prey could impair the acquisition of expertise by reducing hunting success. At the individual-level, our results show that prey speed shapes how individual predators acquire expertise. Therefore, our study outlines how prey antipredator behaviour can mediate the acquisition of expertise in predator populations.

@@

Predation is a fundamental biological process driving ecosystem functioning. Recent advances in our understanding of the mechanisms (e.g. foraging mode, reciprocal plasticity, size-dependent foraging) driving predator foraging success at the individual and population level have allowed ecologists to predict their consequences on prey behaviour and population dynamics [@Peckarsky.etal2008;@gravel.etal2013;@wirsing.etal2021]. Prey typically respond to predation by managing tradeoffs between vigilance and foraging depending on the type of predation regime [@preisser2007]. Such prey antipredator resonses are hypothesized to drive differences among predators in their hunting success by disrupting the acquisition of hunting expertise [@wooster.etal2023], which is deemed essential for many predators to reach adulthood and survive [@woo.etal2008;@wooster.etal2023].

>>>>

The acquisition of expertise is crucial for predators to be successful hunters. To do so, predators need to hone their skills and acquire knowledge through repeated and extensive practice. However, recent studies suggest that predators may not necessarily increase their success with expertise when prey use antipredator tactics to evade detection and pursuit. Therefore, antipredator tactics may impair the acquisition of predator expertise, but there are limited empirical assessments showing how predators acquire expertise through repeated encounters with their prey due to the challenges of monitoring direct interactions in the wild. Here, we use a virtual predator-prey system (the game *Dead by Daylight*) to investigate how experience shapes individual and population hunting success in human predators across repeated interactions with their prey. We show that predators optimized prey consumption as they gained experience, indicating that they acquired expertise through extensive practice. At the population-level, we found that faster prey impair the acquisition of expertise by reducing hunting success. Prey speed was also an important mediator of this relationship at the individual level, driving differences among predators in the development of expertise. Our study outlines how prey antipredator behaviour can mediate the acquisition of expertise in predator populations.

@@

Predation is a fundamental biological process involved in ecosystem functioning. Recent advances in our understanding of the mechanisms (e.g. foraging mode, reciprocal plasticity, size-dependent foraging) driving predator foraging success at the individual and population level have allowed ecologists to predict their consequences on prey behaviour and population dynamics [@Peckarsky.etal2008;@gravel.etal2013;@wirsing.etal2021]. Prey typically respond to predation by managing tradeoffs between vigilance and foraging depending on the type of predation regime [@preisser2007]. Such prey antipredator resonses are hypothesized to drive differences among predators in their hunting success by disrupting the acquisition of hunting expertise [@wooster.etal2023], which is deemed essential for many predators to reach adulthood and survive [@woo.etal2008;@wooster.etal2023].
